Sediment transport capacity response to variations in water discharge in pressurized subglacial channels

### 2025-02-03 revised_v1/response_R1.tex

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}

\vspace{.1cm}

Best regards,

\vspace{.1cm}

Ian Delaney on behalf of all authors

\textbf{General Comments}

\begin{itemize}

\item \textbf{In my view the paper is interesting and relevant, and hence worthy of publication. I have a couple
    of general observations, followed by a list of minor suggestions.
    My general observation relates to the structure of the main findings reported by the manuscript.
    In reading the Results and Discussion I was sometimes lost as to what I was learning that was new.
    I think this was mainly because the structure lacked a clear and logical progression from one
    general finding through progressively more specific findings. This may well stem an initial
    imperfect expression of the objectives.}

  \textit{We thank the Reviewer for the positive assessment of our work and take note of the structural issues along with the organizational ones. }

\item  \textbf{Here, most of what is novel seems to fgall under Objective
    2 and not under Objective 1. Indeed, isn’t Objective 1 (‘to establish whether sub-seasonal water
    discharge can co-vary with sediment transport capacity in subglacial systems’) already established
    beyond reasonable doubt? To me, and I believe the introduction to this manuscript, it clearly
    ‘can’… (note the objective refers to ‘can it’, and not to ‘does it’ [to which the answer is also almost
    certainly ‘yes’ but perhaps with a little more room for maneuver]). I would recast the objectives
    to reflect better the new material presented in this manuscript (closer to objective 2 in fact, which
    I feel could readily be sub-divided). Having considered this, I feel the progression of the Results
    could be far more accessible to the reader and allow the manuscript to focus on a few really key
    points (which I think are currently slightly lost in the density of the presentation).
    Interpretation is presented within each Results section (e.g., see ‘due to… on L174). This
    continues throughout Results, and I would separate all of the explanation and hence
    interpretation out form the results.}

  \textit{We value this comment greatly. Upon consideration, we have chosen to rephrase  objective one as ``to establish the hydrological conditions where sediment transport capacity co-varies
    with water discharge in subglacial systems.'' 
    In our view this objective focus the attention on the differences between the ALPINE and ICESHEET scenarios. It also is practical when evaluating the differences in smooth of water discharge, as discussed in the next point.
    We appreciate the need to remove interpretation from the results section. We have done this. Additionally, Results Section 4.3 with the algebraic formulations has been moved to the discussion, as we think the pipe flow and steady state R-channel scenarios serve as useful end members through which to discuss the results.
  }

\item \textbf{There are some very important points in Results, which I would attach some key data to
    and make more explicit here and in the Conclusion and Abstract. I’m thinking for example of
    associating the (increased) hysteresis as a function of discharge with a headline \% change in
    sediment transport capacity. E.g., “thus, a more variable, but typical, Alpine Q record can
    transport up to *\% greater sediment per cumec than a less variable ice sheet-type Q” and/or “the
    greater variability in Q from Alpine than from ice sheet style Q can account by an offset of peak
    sediment discharge from peak meltwater discharge by up to * \% of the cycle involved”. At present,
    I feel the manuscript lacks this incisive illustration to bring the scale of the results home to the
    reader.}

  \textit{This is an excellent comment. We have worked to address this need for a clearer hysteresis explanation.
    One challenge with the approach recommended by the reviewer is that sediment transport capacity scales very non-linearly with water discharge, especially in subglacial systems. This is shown in Section 4.3.
    As a result, findings across hydrographs would not really be comparable even if they were scaled with a discharge quantity.
    Instead, we will add several paragraphs that to section 4.1 that discuss The role of water discharge variability on hysteresis.

  \textit{Here the aim is to evaluate the effects of smoothing hydrographs, yielding insights in to the hydrological conditions where discharge and sediment transport capacity can co-vary.
    Note that the smoothed hydrographs used for the analysis here are presented in the supplement. 
    A version of the new paragraphs is as follows.}

Water discharge and sediment transport characteristics correlate better in \icesheet{} with lower diurnal variations in water discharge compared to the \alpine{}.
To evaluate the impact of water discharge variability, we smoothed the two hydrographs across different periods and evaluated the Spearman rank correlation between sediment transport characteristics and water discharge.
We assume that a higher rank correlation indicates a reduced amount of hysteresis and behavior more similar subaerial channels (Figure~3).

Smoothing water discharge over periods longer than one day causes a substantial increase in rank correlation (Figure~5).
This increase occurs as diurnal variations are removed (Figures S1-S14).
Correlations in sediment transport characteristics in \icesheet{} remain higher than \alpine, which has more water discharge variations even when smoothed compared to \icesheet{} (Figure~S1-S14).

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.8\linewidth]{../Fig4.png}
  \caption{Spearman rank correlation between water discharge over a smoothed period and sediment transport characteristics (a) water velocity, (b) shear stress, (c) width-integrated shear stress. We assume that higher correlation scores approach subaerial behavior.
  }
  \label{fig:corri}
\end{figure}

\item \textbf{Less importantly, but something I’d like to check, is that the fundamental conclusions are
    applicable more generally. Since they derive from modelling based on two ‘type’ datasets, there
    may be a possibility that not all key findings hold for certain other variations of subglacial
    discharge pattern. I have every reason to believe they do, but I feel the manuscript should report
    that this has been done more widely and the key results/patterns still hold. I don’t think there’s
    any need to include any such results into a revised manuscript – but I would just make the
    statement that while, for simplicity of illustration, two typical but contrasting datasets are
    explored here, the model has been run with other data sets and the key do hold more generally.}

  \textit{ We thank the reviewer for this comment as it, in part, encouraged us to add the additional section above with the water discharge smoothing.
    The other hydrological scenario we are aware of not explicitly considered here is Antarctica, where little surface melt occurs.
    Thus, we expect variations in water discharge to be minimal.
    The following sentence has been added to the conclusion:}
    ``Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size, such as in Antarctica with minimal surface melt input.''

\item   \textbf{
    Incidentally, I would also include an explicit disclaimer early on that the analysis only relates to
    pressurized subglacial flow and not to open-channel flow prominent approaching the terminus of
    many valley glaciers etc.}

   \textit{ We will add the following sentence to the model implementation section:}
``We also note that the subglacial results apply to pressurize subglacial channels, not depressurized one that can occur below glacier termini \citep{perolo2018}''
\end{itemize}

\textbf{Specific Comments}

\begin{itemize}
\item \textbf{Title:}The title states the obvious from prior research and I would recast it to
  reflect more accurately the new and original contribution of this
  manuscript. Also, if the authors elect to not change the title I would change
  the existing wording to ‘transport capacities respond differently’.
  Grammatically, this is not clear given the two capacities noted earlier in the
  title, but stylistically I think ‘capacities respond’ reads better than does
  ‘capacity respond’. (I would change the whole thing anyway… Why not start
  by considering ‘Sediment transport capacity response to variations in water
  discharge in pressurized subglacial channels’)

  \textit{Excellent suggestion, title follows the reviewer's suggestion: Sediment transport capacity response to variations in water discharge in pressurized subglacial channels}

\item \textbf{3:} ‘full’ subglacial channels?

  \textit{Done.}

\item \textbf{5:} There is no need to mention ‘over days’ here (indeed, it is misleading
  anyway without quantification of how much change as a function of time).
  I would just leave it as changes slowly (or slowly relatively to variations in
  water Q).

  \textit{Done.}

\item \textbf{7:} ‘Sheet’ is missing (the authors really should have picked up on this obvious
  typo)

  \textit{Done.}

\item \textbf{8:} This hysteresis causes (no need to qualify the type of hysteresis since the
  use of ‘this’ serves the purpose)

  \textit{Done.}

\item \textbf{64:} I think there is earlier use of transport-limited (in which case the definition
  should be presented there, and not here)

  \textit{We thank the reviewer for the careful read. The definition was presented in Line 28, and thus we will remove the definition here.}

\item \textbf{75-76 and 80-81} I’d remove summary of the results of this study from the Introduction. They
  are not yet established.

  \textit{These lines will be removed.}

\item \textbf{70-71} In phrasing objective 1 (in more than one place in the manuscript),
  shouldn’t the dependent variable come first? Thus, I would write it as ‘to
  establish whether sediment transport capacity can co-vary with water
  discharge in subglacial systems’.

  \textit{Given the discussion above, the text will read: ``to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems''}

\item \textbf{79} Given the apparent relevance of the Alley et al. study, I feel it’s key findings
  should be presented in detail here to provide more direction to the specific
  objectives of this study.

  \textit{ We believe that this is discussed in Lines 51--54 in the introduction.}

\item \textbf{159} The parameter values are not purely random (but random within certain
  sensible ranges, no?).

  \textit{The text will now read: ``with random parameter values in a range''}

\item \textbf{171-} I get a little muddled right at the start here – not helped by the sub-heading,
  which I would recast as “Influence of subglacial channel size on the timing
  and variability of…”. The section (and others following) also combines
  results with explanation (i.e., interpretation) which I feel also adds
  unnecessarily to the difficulty in following the progression. I would separate
  results from interpretation.

  \textit{This section will be retitled:} ``Timing of subaerial and subglacial sediment transport capacity variations.''
  \textit{This discusses the role water discharge on hysteresis in water velocity or sediment transport capacity. Thus dealing with objective 1 of the manuscript. We have also noted the comment about results and interpretation. These have been removed from the manuscript}

\item \textbf{172} This narrative jumps straight in with the ambition to ‘quantify the sources
  of increased variability…’ but the nature of that increased variability has not
  yet be established or illustrated. This really needs building up more
  progressively – bringing the reader through with the first-order model
  results. Then go on to explore more and more detailed influences and
  relationships. Most of this is, in fact, in Figure 2 – but the reader is not
  guided through the fundamental relationships here.

\ian{Comeback here}

\item \textbf{173} Change rather than evolution? (elsewhere too)

  \textit{We appreciate the comment. We believe that given the nature of subglacial channels as describe in Equation 3, evolution is a more accurate description. To us, ``change'' could include a modification for any number of reasons. However,
    ``evolution'' conotates the dependence on anticedent conditions which we believe to be more accurate in this case. }

\item \textbf{Fig 2 caption} Is it really an arbitrary y axis scale? If it is truly arbitrary, does it even need
  a scale range?

  \textit{We believe that description of the insets' axis is needed as these lines have the same axis in the main plots. However, their very different values mean that in makes sense to adjust values to examine their changes during the extreme events. The text has been modified slightly to read: ``different y-axis ranges for subaerial and subglacial values.''}

\item \textbf{179-183} Is this not already established? If so, it should be presented clearly in the
  Introduction and developed as necessary here. It is also interpretation.

  \textit{ To the best of our knowledge this is the first time that this finding is being presented.
    We would welcome any citations or references that suggest otherwise. We believe that the content in this paragraph describes the response to water discharge in Figure 2. We agree that the reference to the Methods section, gave the impression of the interpretation. The paragraph will be adjusted slightly to read:}

    Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~2).
    In the subglacial channel, peaks in model outputs generally occur when water discharge increases, but before the maximum water discharge.
    As the water discharge stabilizes at its peak, channel growth continues (Figure~2 a, e), causing water velocity and other model outputs to decrease from their peak values (Figure~2 b-d, f-h).
    Subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.''}

\item \textbf{Fig 3 (and Fig 4)} Shouldn’t the dependent variable be plotted on y and the independent
  variable on x? I for one at least think this way around and therefore find
  these plots a little confusing.
  Is there a need for the axis labels to be at an angle rather than parallel to
  the axis? If not, I would reorientate them parallel in all cases.

  \textit{The axes will be changed in both figures}

\item \textbf{193-4} This is interpretation.

  \textit{This sentence will be removed.}

\item \textbf{195} The reference to ‘increased variability’ needs to make it clear what property
  is being referred to. Like the previous sub-heading, this sub-heading is a
  statement of findings before those results have been presented. I would
  avoid this and present it as the less assuming: ‘Influence of…’

  \textit{The section head will be modified to read:} Variability in sediment transport capacity across a range of channel shapes, slopes, and friction values

\item \textbf{205-6} Interpretation in Results

  \textit{Removed.}

\item \textbf{210-11} Interpretation in Results

  \textit{ Sentence will be changed to:''Smaller values of channel factor $\beta$, creating low and broad channels where the channel width grows more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~4).
''}

\item \textbf{220 Typo (subglacial)}

  \textit{Fixed.}

\item \textbf{Section 4.3} A lot of this reads more like methods and background modelling operation
  and/or equation terms than Results of the modelling. If it is all results then
  the key findings are difficult to pull out from the density of the presentation
  – perhaps some of it could be omitted or moved to supplementary?

  \textit{We have considered the comments here, and with the editors support, think it fits best here.
    In a previous version of the manuscript, it was included in the supplementary material. However, after those reviews, we believe that some misunderstandings emerged. We believe that Tables 2 and 3 in the section are valuable results in evaluating the sensitivity of sediment transport capacity to water discharge under different assumptions.
    The importance of these findings are discussed in lines xxx yyy.
  }

  \textit{We will modify the introductory text to read: ``The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~4.3). 
Additionally, results demonstrate the impact of water discharge variability on sediment transport capacity.
Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient.''
}

\item \textbf{Fig 5} I find the figure too small. If larger the font would be easier to read and the
  y axes could have the property written out rather than represented by
  symbols. I’m sure there is no need for it to be this small.

  \textit{We will increase the size of the axis labels.}

\item \textbf{232} Typo? (‘typesc’)
  \textit{Removed.}

\item \textbf{235} The text: ‘…as used as proxy for…’) doesn’t make sense (or at least could be
  worded more clearly).

\textit{Text will read: ''additionally width-integrated shear stress is assessed, instead of sediment transport, as above.''}

\item \textbf{284} (maybe
  elsewhere too)
  The wording should be “greater …. than” and not ‘greater… compared
  with…).

  \textit{Done.}

\item \textbf{297} “…itself dependent of…” doesn’t make sense to me

  \textit{Text now reads: ''which is dependent''}

\item \textbf{298} I don’t believe there is sufficient evidence to make the claim: ‘While
  transport limited states likely do not occur at many glaciers’. In my own
  experience most glacier do have at least parts of the bed that are sediment-
  rich and hence have the capacity to be transport limited.

  \textit{We agree with  about the lack of data. This sentence will been removed. }

\item \textbf{302-4} I would try to pull some key quantification out here and present it as a
  headline figure to illustrate the magnitude of the effect.

\item \textbf{319} I’m not sure ‘sporadic’ is the right word here. Please check you mean this.

  \textit{We will replace ``sporadic'' with ``variable''.}

\item \textbf{321-22} I see the point being made here but it’s not black or white. Surely there is
  some control exerted by discharge; it’s just that it is complicated by the
  additional forces in the subglacial scenario. The way this is presented here
  may be technically ok but the reader could be forgiven for coming away
  with the feeling that there is no solid relationship between Q and sediment
  transport capacity is these cases – which would not be accurate. I feel it
  would be particularly helpful if the effect could be quantified in some
  generalized or headline way in terms of the likely (or maximum if you
  prefer) effect on a typical alpine and/or ice sheet hydrograph. That figure
  could then also go into the Conclusions and Abstract.

  \textit{The sentence will read: ``This hysteresis can limit water discharge as an indicator of sediment discharge capacity in these systems, especially when water discharge is highly variable and out of equilibrium with subglacial channel size.''}

\item \textbf{394} Surely the relationship can be characterized a little more usefully than
  simply to state that it is ‘incoherent’. This also relates to my point above
  relating to lines 321-22. Can this be characterized accurately or
  systematically so the reader has a little more to learn in terms of the nature
  and magnitude of the effects?

### 2025-02-03 Delaneyetal2024.tex
small change

%%Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

@@

\title{Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations}

@@

\begin{abstract}
 Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
  In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
  However, in subglacial channels, water is pressurized by the ice above, and they grow in response to frictional heating of water flowing through them. 
  As a result, water discharge changes mainly result in velocity variations, as the channel geometry evolves slowly (over days).
  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice. 
  Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
  This hysteresis in channel size causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
  A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
  Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge. 
  The implications of these findings are discussed in the context of sediment discharge from glaciers with different hydro-climatic forcings.
  We also discuss the impact of different assumptions of channel behavior on sediment transport capacity.
  These findings can improve the interpretation of sediment discharge records in glacierized catchments.
\end{abstract}

@@

In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, which is defined as the amount of sediment the water can carry.
In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows over \citep{shields1936,meyer1948,engelund1967}, along with the width of the channel bottom $w$ over which sediment mobilizes.
The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau}
    \tau \propto v^2.
  \end{equation}
\end{linenomath*}
% 
Following mass conservation, the mean velocity of the water flowing through a  channel is
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area.

@@

Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment access or production.
Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, where sediment transport responds to the channel's hydraulic conditions, instead of sediment availability.
The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchments \citep[e.g.][]{ganti2016,mancini2023}.
This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019b} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.

This manuscript has two objectives: 1)  to establish whether sub-seasonal water discharge can co-vary with sediment transport capacity in subglacial systems, and 2) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones.
We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
Model outputs demonstrate the specific processes that can drive variability in sediment discharge capacity in subglacial systems.
We also run a model ensemble to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial channels.
Lastly, we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, steady-state R-channel, and pipe flow conditions.
These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel assumptions.
Findings indicate differences in the relationships amongst water discharge, channel geometry, water velocity and sediment transport capacity in subglacial and subaerial channels.
The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.

@@

The subglacial channel model accounts for the channel geometry and the water's velocity to evaluate the shear stress of water flowing across sediments underneath a glacier.
We use a lumped hydraulics model from \citet{werder2010b}, itself based upon  \citet{clarke1996}.

Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean thickness of $h_{ice}$.
The channel size grows from melt due to frictional heating from water flow and closes due to ice creep.
The formulation here does not consider the englacial storage of water.
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\Delta h$ is the hydraulic head drop change over $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.

@@

Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
Thus, the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).

\begin{table}[hbt!]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier.
    A second line refers to the range of values examined in the parameter search.}
  \small 
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\
         && (\alpine{} or \icesheet{})\\
    \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (subglacial, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (subglacial, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Channel depth (subaerial) & $H$&& $\mathrm{m}$\\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (subglacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\
    Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\

>>>>

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

@@

\title{Sediment transport capacity response to variations in water discharge in pressurized subglacial channels}

@@

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with water velocity and channel width.
  In full subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
  However, in subglacial channels, water is pressurized by the ice above, and they grow in response to frictional heating of water flowing through them. 
  As a result, rapid changes in water discharge  mainly result in velocity variations, as the channel geometry evolves slowly.
  Here, we present formulations of sediment transport capacity in different channel types and apply subglacial and subaerial hydraulics models to hydrographs from an Alpine glacier and the Greenland Ice Sheet. 
  Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
  This hysteresis causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
  Reducing water discharge variability lessens the hysteresis effects, in some cases to the point where a covarying relationship between water discharge and sediment transport capacity may occur, similar to subaerial systems.
  A second set of numerical experiments shows that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
  Yet, results and  formulations of subglacial sediment transport capacity show that its variability can approach that of subaerial systems when subglacial channel size is in equilibrium with water discharge.
  We present these results in the context of different formulations of the relationship between water discharge and sediment transport capacity based assumptions of subglacial and subaerial flow regimes.
  The implications of these findings help to evaluated sediment discharge from glaciers with different hydro-climatic forcings.
  These findings can improve the interpretation of sediment discharge records in glacierized catchments.
\end{abstract}

@@

In a transport-limited regime, sediment discharge is controlled by sediment transport capacity, which is defined as the amount of sediment the water can carry.
In both subglacial and subaerial channels, sediment transport capacity depends on the shear stress between water and the sediment it flows over \citep{shields1936,meyer1948,engelund1967}, along with the width of the channel bottom $w$ over which sediment mobilizes.
The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau}
    \tau \propto v^2.
  \end{equation}
\end{linenomath*}
% 
The mean velocity of the water flowing through a  channel is
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area.

@@

Despite the different physical processes between the subaerial and subglacial systems, observed water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
Many of these studies discuss the disparity between water discharge and sediment export, especially given the role of sediment access or production.
Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited regime, instead of sediment availability.
The response of sediment transport capacity to water discharge variations may affect the interpretation of sediment transport records in these glacierized catchments \citep[e.g.][]{ganti2016,mancini2023}.
This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019b} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.

This manuscript has two objectives: 1)  to establish the hydrological conditions where  sediment transport capacity co-varies with water discharge in subglacial systems,
and 2) to evaluate the variability of  sediment transport capacity in subglacial channels with evolving channel size and different hydrographs compared to subaerial ones.
We use numerical models to examine sediment transport capacity in both subglacial channels that evolve in size and subaerial channels that remain fixed.
They are applied to proglacial hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and a land-terminating glacier in Greenland (Leverett Glacier).
We also run a model ensemble to evaluate the sediment transport capacity variability in subglacial channels compared to subaerial channels.
Lastly, we present algebraic formulations of the sediment transport capacity response to water discharge in subaerial, steady-state R-channel, and pipe flow conditions.
These formulations extend the relationships presented in \citet{alley1997} and illustrate the sediment transport capacity behavior of different channel assumptions.
The manuscript then discusses the implications of variability in sediment transport capacity from glaciers and the interpretation of sediment transport records.

@@

The subglacial channel model accounts for the channel geometry and the water's velocity to evaluate the shear stress of water flowing across sediments underneath a glacier.
We use a lumped hydraulics model from \citet{werder2010b}, itself based upon that of \citet{clarke1996}.

Here, it is assumed that the water is transported through a subglacial channel \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972} beneath a glacier with channel length $l$, with a flat bed and a mean ice thickness of $h_{ice}$.
The channel size grows from melt due to frictional heating from water flow and closes due to ice creep.
The formulation here does not consider the englacial storage of water.
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\overline{h}\right)^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~\ref{table:vpm}), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\overline{h}=\frac{1}{2}(h+h_p)$ is the mean hydraulic head in the channel with $h_p$ being the proglacial hydraulic head equal to zero, $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.

@@

% Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
% Thus, the channel size in the subaerial model has no history dependence on the discharge $Q_w$, whereas the subglacial one does (Equation~\ref{eq:dS_dt}).

\begin{table}[hbt!]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier.
    A second line refers to the range of values examined in the parameter search.}
  \footnotesize
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\
         && (\alpine{} or \icesheet{})\\
    \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (subglacial, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (subglacial, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Channel depth (subaerial) & $H$&& $\mathrm{m}$\\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (subglacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Mean hydraulic head & $\overline{h}$&& $\mathrm{m}$\\
    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\
    Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\
